---
title: "Example T-F0 R code"
author: "Pieter A. Arnold"
output:
  pdf_document: default
  html_notebook: default
---


Example two-thermocouple hot or cold Tcrit, T50, Tmax extracting code
This code is designed to extract Tcrit, T50, Tmax values from Imaging fluorometer data, and datalogger data with two thermocouples used as representative measures of leaf temperatures on samples on a Peltier plate, and a reference ID file. The example uses 24 leaf samples from three species using a dry substrate, 30 degrees C per hour ramp rate and a heat tolerance assay ramping from 20 to 65 degrees C. Refer to the manuscript for biological and methological details.  

```{r load libraries}

library(chron)
library(segmented)
library(ggplot2)
library(scales)

```


``` {r vlookup function}

vlookup <- function(ref, table, column, range = FALSE, larger = FALSE)
{
  if(!is.numeric(column) & !column %in% colnames(table)) {
    stop(paste("can't find column",column,"in table"))
  }
  if(range) {
    if(!is.numeric(table[,1])) {
      stop(paste("The first column of table must be numeric when using range lookup"))
    }
    table <- table[order(table[,1]),] 
    index <- findInterval(ref,table[,1])
    if(larger) {
      index <- ifelse(ref %in% table[,1],index,index+1)
    }
    output <- table[index,column]
    output[!index <= dim(table)[1]] <- NA
    
  } else {
    output <- table[match(ref,table[,1]),column]
    output[!ref %in% table[,1]] <- NA 
  }
  dim(output) <- dim(ref)
  output
}

```


```{r folders files and meta-data}

# Set folder directory where files are
folder    <- "/Users/pieterarnold/Desktop/Example T-F0 R code/"
i_files   <- list.files(folder, pattern = "*ID.csv", 
                        full.names = TRUE, recursive = TRUE)
f_files   <- list.files(folder, pattern = "*fluorescence.csv", 
                        full.names = TRUE, recursive = TRUE)
t_files   <- list.files(folder, pattern = "*thermocouple.csv", 
                        full.names = TRUE, recursive = TRUE)

# Check the number of fluorescence, thermocouple, and ID files are the same
length(f_files); length(t_files); length(i_files)

# Set meta-data information: ADJUST AS NEEDED
date     <- 20190808 # the date of the assay
n_run    <- 1 # the assay / run identification number
nsamples <- 24 # number of leaf samples / areas of interest (AOIs)
rate     <- 30 # ramp rate

```


```{r Reading normalising and subsetting data}

# Read in the imaging fluorometer data file (with timestamp and F0 values only)   
fluor  <- read.csv(f_files)
head(fluor)
str(fluor)

# Read in the thermocouple data file 
# (with timestamp and thermocouple temperature measurements)
therm  <- read.csv(t_files)
head(therm)
str(therm)

# Set the data output file matrix (n calculated parameters wide x n samples long)
output <- data.frame(matrix(NA, nsamples, 7), stringsAsFactors = FALSE)
  
# Have to make time values characters in order to make them times. 
# Time may be read as factor automatically.
fluor$PAMtime <- as.character(fluor$PAMtime)
therm$Tctime  <- as.character(therm$Tctime)
# Note below will not work unless times in csv are already formatted to hh:mm:ss
fluor$PAMtime <- chron(times = fluor$PAMtime)
therm$Tctime  <- chron(times = therm$Tctime)
#fluor <- fluor[,colSums(is.na(fluor))<nrow(fluor)]
#therm <- therm[,colSums(is.na(therm))<nrow(therm)]

# Use the function 'vlookup' to match the thermocouples temperatures and 
# the fluorescence values by matching the times
for (j in 1:nrow(fluor)){
  fluor$Temp1[j]     <- vlookup(fluor$PAMtime[j], therm, 2, TRUE)
  fluor$Temp2[j]     <- vlookup(fluor$PAMtime[j], therm, 3, TRUE)
  fluor$Tempmean[j]  <- vlookup(fluor$PAMtime[j], therm, 4, TRUE)
}
  
# Subset the data to set sensible limit within which the Tcrit value should be 
# (works for hot and cold tolerance assays)
xlow  <- ifelse(read.csv(i_files)$Temp[1]=="Hot", 30, -20)
xhigh <- ifelse(read.csv(i_files)$Temp[1]=="Hot", 65, -2)
  
# Determine how many degrees C before the 50% fluorescence to fit the regression within: 
# this can be optimised by adjusting based on the ramp rate that is being used 
x_postbuffer <- 1 # parameter to change how many degrees C after 
                  # 50% fluorescence to fit the regression within
if(xlow=="30") { 
  x_prebuffer <- 12 } else if(rate >= 60) { 
    x_prebuffer <- 12 } else if(rate >= 30 & rate < 60) { 
      x_prebuffer <- 5 } else if(rate >= 15 & rate < 30) { 
        x_prebuffer <- 3 } else  { 
          x_prebuffer <- 2 }

# Subset the fluorescence dataset to be within the xlow and xhigh values 
# dependent on the direction of T-F0 assay
if(xlow=="30") {
  fluor <- subset(fluor, fluor$Tempmean > xlow + 1 & fluor$Tempmean < 59) } else { 
    fluor <- subset(fluor, fluor$Tempmean < xhigh & fluor$Tempmean > -20) }
  
# Plot the datalogger temperature against time
# png(filename = paste0(date[i], "_run", n_run[i], "_thermocouples.png"), 
# width = 700, height = 700, units = "px")

par(mfrow = c(1,1), mar = c(4,4,1,1))
if(xlow=="30") {
  plot(Temp1 ~ PAMtime, data = fluor, type = "l", col = "blue", xlab = "Time", 
       ylab = "Temperature", ylim = c(30, 60))
} else {
  plot(Temp1 ~ PAMtime, data = fluor, type = "l", col = "blue", xlab = "Time", 
       ylab = "Temperature", ylim = c(-20, 0))
}
points(Temp2 ~ PAMtime, data = fluor, type = "l", col = "green") 
points(Tempmean ~ PAMtime, data = fluor, type = "l", col = "red")
legend("right", legend = c("Thermocouple 1", "Thermocouple 2", "Thermocouple mean"), 
       col = c("blue", "green", "red"), lty = 1, bty = "n")
#dev.off()
  
# Safeguard for if there is an electrical glitch for the fluorescence signal
# Prior to rescaling, remove that row of data
minsignal <- which.min(fluor$F1)
fluor     <- fluor[-c(minsignal),]

# Normalise values for each sample so that the first value is taken as the lowest value 
# (i.e., later values cannot be below the starting value, so the first value is scaled 
# to zero, which standardises 50% and 100% fluorescence values).

fluor1 <- NA
fluor2 <- data.frame(matrix(NA, 200, nsamples), stringsAsFactors = FALSE)
for (column in 1:(nsamples)) {
  fluor1[column] <- fluor[column + 1]
  #l <- lengths(fluor1[column])
  xd <- as.vector(fluor1[column][[1]])
  xdr <- as.data.frame(rescale(xd, to = c(0,1), from = range(xd, na.rm = TRUE)))
  xdr[nrow(xdr):200,] <- NA
  fluor2[,column] <- as.vector(xdr)
}

# Recompile the dataframe that has had F0 values normalised F__ columns names are now 
# named as X__ and the dataframe used below is called fluorN
fluor[nrow(fluor):200,] <- NA
fluorN <- (cbind(fluor, fluor2))
  

```

``` {r Extract Tcrit (CTmin orCTmax) T50 Tmax}

# for loop for each rescaled sample to calculate the Tcrit, T50, Tmax. 
for (m in ((nsamples + 5):ncol(fluorN))) { 
  # The "nsamples + 5" sets which column is the first of the normalised F0 values 
  # png(filename = paste0(date[i], "_run", n_run[i], "_Tcrits", m-nsamples[i]-5, ".png"), 
  # width = 700, height = 700, units = "px")
  tryCatch({
    par(mfrow = c(1,1), mar = c(4,4,1,1))
    plot(fluorN[,m] ~ Tempmean, data = fluorN, type = "l", ylab = names(fluorN[m]),
         xlim = c(xlow, xhigh), ylim = c(0,1))
    maxval    <- which.max(fluorN[,m])
    tempatmax <- fluorN$Tempmean[maxval]
    
    # If/else to make the extraction process work for both cold or hot tolerance assay data  
    if(xlow=="30") {
      val50 <- which(abs(subset(fluorN, fluorN$Tempmean < tempatmax)[,m]-0.5)==
                       min(abs(subset(fluorN, fluorN$Tempmean < tempatmax)[,m]-0.5)))
      tempat50  <- fluorN$Tempmean[val50][1]
      thresholdval <- which(fluorN[,m] >= fluorN[val50[1],m])
      tempatthreshold <- fluorN$Tempmean[thresholdval[1]] + x_postbuffer
    } else {
      val50 <- which(abs(subset(fluorN, fluorN$Tempmean > tempatmax)[,m]-0.5)==
                       min(abs(subset(fluorN, fluorN$Tempmean > tempatmax)[,m]-0.5)))
      tempat50  <- fluorN$Tempmean[val50][1]
      thresholdval <- which(fluorN[,m] > fluorN[val50[1],m])
      tempatthreshold <- fluorN$Tempmean[thresholdval[1]] - x_postbuffer
    }      
      
    # Display the subset range for searching for the Tcrits 
    if(xlow=="30") {
      segments(tempat50 - x_prebuffer, 0, tempat50 - x_prebuffer, 1, lwd = 2, 
               lty = 3, col = "blue") } else {
      segments(tempat50 + x_prebuffer, 0, tempat50 + x_prebuffer, 1, lwd = 2, 
               lty = 3, col = "blue")
    }
    segments(tempatthreshold, 0, tempatthreshold, 1, lwd = 2, lty = 3, col = "blue")
    
    # Subset the data within which to apply the regression to find the Tcrit value  
    if(xlow=="30") {
      fluor_sub <- subset(fluorN, fluorN$Tempmean > xlow & 
                            fluorN$Tempmean < tempatthreshold[1] & 
                            fluorN$Tempmean > (tempat50[1] - x_prebuffer)) } else {
                            fluor_sub <- subset(fluorN, fluorN$Tempmean > xlow & 
                                                  fluorN$Tempmean > tempatthreshold[1]
                            & fluorN$Tempmean < (tempat50[1] + x_prebuffer))
                          }
    # Run the breakpoint regression and add the points for Tcrit, T50, Tmax to the plot
    response   <- fluor_sub[,m]
    model3     <- lm(response ~ Tempmean, data = fluor_sub)
    seg_model3 <- segmented(model3, seg.Z = ~ Tempmean, npsi = 1, data = fluor_sub)
    fitted_val3 <- fitted(seg_model3)
    breakmodel3 <- data.frame(Temperature = fluor_sub$Tempmean, fluor_sub = fitted_val3)
    tcritmean     <- round(seg_model3$psi[[2]], 2)
    tcriterrmean  <- round(seg_model3$psi[[3]], 2)
    tcrittextmean <- paste("Tcrit-break =", tcritmean, "°C")
    lines(fluorN[,m] ~ Tempmean, data = fluorN, lwd = 4)
    lines(fluor_sub ~ Temperature, data = breakmodel3, type = "l", col = "red", lwd = 4)
    points(fluor_sub[which.min(abs(fluor_sub$Tempmean - tcritmean)),m] ~ tcritmean, 
           pch = 24, bg = "orange", cex = 2)
    points(fluorN[maxval,m] ~ tempatmax[1], pch = 21, bg = "red", cex = 2)
    points(fluorN[val50[1],m] ~ tempat50[1], pch = 21, bg = "green", cex = 2)
    tempatmaxtext <- paste("Tmax =", round(tempatmax, 2), "°C")
    tempat50text  <- paste("T50 = ", round(tempat50, 2), "°C")
    legend("right", legend = c(tempatmaxtext, tempat50text, tcrittextmean), 
           pt.bg = c("red", "green", "orange"), pch = c(21, 21, 24), bty = "n")
    
    # Fill out the output dataframe to store the Tcrit values 
    k <- m-nsamples-4
    output[k,1] <- date
    output[k,2] <- n_run
    output[k,3] <- names(fluorN)[m]
    output[k,4] <- tcritmean
    output[k,5] <- tcriterrmean
    output[k,6] <- tempat50[1]
    output[k,7] <- tempatmax[1]
      
    }, error = function(e) {cat("ERROR :", conditionMessage(e), "\n")})
    #dev.off()
  }
  
colnames(output) <- c("Date", "Run", "AOI", "Tcrit-break", 
                      "Tcrit-error", "tempat50", "tempatmax")
output
  
# Append the Tcrit output to the ID dataset
id_tcrit_dataset <- cbind(read.csv(i_files), output)
id_tcrit_dataset
# Option to write the dataset as a .csv file to be called into another script
# write.csv(id_tcrit_dataset, paste0(date, "_run", n_run, "_dataset.csv"))

#### END ####
```


